---
title: "Descriptive Statistics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(conflicted)
library(targets)
library(data.table)
library(dplyr)
library(ggplot2)
```

```{r load_data}
tar_load(harmonized_names)
cand_data <- fread("./data/consulta_cand_2020_BRASIL.csv.gz", encoding = "Latin-1")
```

## Coverage of Parents Data

Currently, we have extracted the parents data for mayoral candidates from the 2020 elections. The percentage of candidates with parents data can be found in the following table.

```{r coverage_parents}

```{r coverage}
cand_data <- cand_data[DS_CARGO == "PREFEITO" &
        DS_SITUACAO_CANDIDATO_PLEITO %in% c("DEFERIDO", "DEFERIDO COM RECURSO") &
        DS_SIT_TOT_TURNO != "#NULO#"]

cand_data[, found := fifelse(as.character(SQ_CANDIDATO) %in% harmonized_names$cand_id, TRUE, FALSE)]
cand_data[, found_2 := fifelse(as.character(SQ_CANDIDATO) %in%
        harmonized_names$cand_id[harmonized_names$n_parents >= 2], TRUE, FALSE)]
prop_found_at_least1 <- cand_data[, .(pct_found = 100 * mean(found)), by = "DS_SIT_TOT_TURNO"]
prop_found_at_least2 <- cand_data[, .(pct_found2 = 100 * mean(found_2)), by = "DS_SIT_TOT_TURNO"]

prop_found <- merge(prop_found_at_least1, prop_found_at_least2)
prop_found[
        ,
        DS_SIT_TOT_TURNO :=
                fcase(
                        DS_SIT_TOT_TURNO == "2º TURNO", "2nd Round",
                        DS_SIT_TOT_TURNO == "ELEITO", "Elected",
                        DS_SIT_TOT_TURNO == "NÃO ELEITO", "Not Elected"
                )
]
names(prop_found) <- c("Candidate Status", "at least 1 parent", "2 parents")

prop_found_uf <- cand_data[, .(pct_found = 100 * mean(found)), by = c("SG_UF")]
prop_found_uf[, SG_UF := forcats::fct_reorder(SG_UF, pct_found)]

prop_found2_uf <- cand_data[, .(pct_found = 100 * mean(found_2)), by = c("SG_UF")]
prop_found2_uf <- prop_found2_uf[order(pct_found)]
```


```{r coverage_table}

knitr::kable(prop_found,
        digits = 1,
        caption = "% of Candidates with Parents Data"
)
```

The percentage of candidates with parents data by state can be found in the following plot.
```{r coverage_plot}
ggplot(prop_found_uf, aes(x = pct_found, y = SG_UF)) +
        geom_point() +
        xlab("% with at Least 1 Parent") +
        ylab("State") +
        cowplot::theme_cowplot() +
        ggtitle("% of Candidates with Parents Data by State")
```



